---
title: "Computing Cumulative Incidence Functions  with the `cshCIF` Package"
author: "Jerome V. Braun"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Computing Cumulative Incidence Functions  with the `cshCIF` Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The `cshCIF` package provides basic functionality for estimating the cumulative incidence function (CIF) for an risk of interest in the presence of competing risks.

## Data example

We will use the [`bmtcrr`](http://www.stat.unipg.it/luca/R/) data set (included preloaded in the package).

```{r}
library(cshCIF)

data(bmtcrr)
```

The `cshCIF` package requires data in a wide format, with separate time and censoring status variables for each cause. 

Also, the current functionality only allows regression style effects, so all categorical variables, transformed variables, and interactions must be recoded by hand in advance.

```{r}
df <- data.frame(
    RTime = bmtcrr$ftime,
    RStatus = ifelse(bmtcrr$Status == 1, 1, 0),
    CTime = bmtcrr$ftime,
    CStatus = ifelse(bmtcrr$Status == 2, 1, 0),
    Age = bmtcrr$Age,
    Sex = ifelse(bmtcrr$Sex == "F", 1, 0),
    CR1 = ifelse(bmtcrr$Phase == "CR1", 1, 0),
    CR2 = ifelse(bmtcrr$Phase == "CR2", 1, 0),
    CR3 = ifelse(bmtcrr$Phase == "CR3", 1, 0),
    Source = ifelse(bmtcrr$Source == "PB", 1, 0)
)
```

## Fitting the model

The input to the model must be specified via a list, with the cause of interest as the first element in the list.

Each element of the list is itself a list, defining a name for the cause, the variable giving the time of the event due to the cause, a status variable defining censoring (1 = event, 0 = censored), and the variables to use in the model for that cause.  In this case, "Relapse" is the event of interest.

We model the risk of relapse as a function of age, sex, phase, and source.  The competing risk is modeled as a function of only age and sex.

First, we use the `csh_censor` function to adjust the censoring to account for a potential follow-up window.  In these data, there is no follow-up window, but the function is run to demonstrate its use anyway.

Next, we fit the cause-specific hazard model.

Finally, we estimate the marginal risk at 50 months for a 30-year old female with phase relapse and source PB + BM along with confidence intervals. 

```{r}
CSH <- list(
    list(Cause = "Relapse", Time = "RTime", Status = "RStatus", vars = c("Age", "Sex", "CR1", "CR2", "CR3", "Source")),
    list(Cause = "Competing", Time = "CTime", Status = "CStatus", vars=c("Age", "Sex"))
)

dfc <- csh_censor(df = df, csh = CSH, fuw = 1)
fit <- csh_fit(dfc, csh = CSH)

Covariates <- data.frame(Age = 30, Sex = 0, CR1 = 0, CR2 = 0, CR3 = 0, Source = 0)

csh_CIF(fit, Covariates, csh = CSH, risk_times = 50, alpha=0.05, method = "loglog")
```

## CIF curves

If we do not specify the risk times of interest, then the risk is computed at the unique times in the data set by default.

The following code calculates the CIF for males and females separately.

```{r}
# Fit the model for PB.

Covariates <- data.frame(Age = 30, Sex = 0, CR1 = 0, CR2 = 0, CR3 = 0, Source = 0)
CIF.0 <- csh_CIF(fit, Covariates, csh = CSH, alpha=0.05, method = "loglog")

# Fit the model for PB + BM.

Covariates <- data.frame(Age = 30, Sex = 1, CR1 = 0, CR2 = 0, CR3 = 0, Source = 0)
CIF.1 <- csh_CIF(fit, Covariates, csh = CSH,  alpha=0.05, method = "loglog")
```

```{r, fig.show='hold'}
library(ggplot2)

ggplot(rbind(CIF.0, CIF.1), aes(x = risk_time, y = Risk)) +
    geom_ribbon(aes(ymin = Risk_LCL, ymax = Risk_UCL), fill = "grey70") +
    geom_step() +
    facet_wrap(~ ifelse(Sex, "Female", "Male")) +
    theme_bw()
```

